---
title: "Introduction to Leaflet"
subtitle: "UF R Meetup"
author: "Justin Millar"
date: "2018/04/18"
output:
  xaringan::moon_reader:
    css: ["theme.css", "fonts.css"]
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---
```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
```

```{r libs, echo=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
library(leaflet)
library(RColorBrewer)
library(sp)
library(raster)
```

class: inverse, center, middle

### So what is Leaflet?

---

# What is Leaflet?

`leaflet` is a package in R that can create interactive maps

--

## What is Leaflet, really?

Leaflet is actually a JavaScript library, and the `leaflet` package is a wrapper that allows us to create Leaflet JavaScript using R code

This isn't really important for using Leaflet in R applications, but may be handy to know for the future

---

```{r install, eval=FALSE}
install.packages("leaflet")
library(leaflet)

leaflet() %>% 
  setView(lat = 29.6436, lng = -82.3549, zoom = 14) %>%
  addTiles() 
```


```{r world, echo=FALSE, out.width='100%'}
library(leaflet)
leaflet() %>% 
  setView(lat = 29.6436, lng = -82.3549, zoom = 14) %>%
  addTiles()
```

---
class: inverse, center, middle

### So kinds of R data can I use?

---

Regular old base R dataframe/matrix with lng/lat columns

Spatial objects from the `sp` package
* SpatialDataframes
* SpatialLines
* Polygons (from shp files)

Map objects from the `maps` package

```{r maps, out.width='100%', eval=FALSE}
fl_counties <- maps::map("county", "florida", 
                         fill = TRUE, plot = FALSE)

leaflet(data = fl_counties) %>% 
  addTiles() %>% 
  addPolygons(fillColor = topo.colors(10, alpha = NULL), 
              weight = 1, color = 'black')  
```

---

```{r maps-show, out.width='100%', out.height='100%', echo=FALSE}
fl_counties <- maps::map("county", "florida", 
                         fill = TRUE, plot = FALSE)

leaflet(data = fl_counties) %>% 
  addTiles() %>% 
  addPolygons(fillColor = topo.colors(10, alpha = NULL), 
              weight = 1, color = 'black')  
```

---

class: inverse, center, middle

### Ok, that's great. But the interaction is more than just zooming in and out, right?

---

### Labels

```{r label, out.width='100%', eval=FALSE}
fl_map <- leaflet(data = fl_counties) %>% 
  addTiles() %>% 
  addPolygons(fillColor = topo.colors(10, alpha = NULL), 
              weight = 1, color = 'black', 
              label = ~as.character(names))
```

--
### Markers

```{r marker, out.width='100%', eval=FALSE}
fl_map %>% 
  addMarkers(lat = 29.64833, lng = -82.34944, 
             popup = "University of Florida")
```

--
### Highlighting

```{r hlgt, out.width='100%', eval=FALSE}
fl_map %>% 
  addPolygons(highlightOptions(color = "red", weight = 3))
```

---
```{r full, out.width='100%', out.height = '100%', echo=FALSE}
leaflet(data = fl_counties) %>% 
  addTiles() %>% 
  addPolygons(fillColor = topo.colors(10, alpha = NULL), 
              weight = 1, color = 'black', label = ~as.character(names), 
              highlightOptions = highlightOptions(color = "red", weight = 3,
                                                  bringToFront = TRUE)) %>% 
    addMarkers(lat = 29.64833, lng = -82.34944, 
             popup = "University of Florida")
  
```

---
class: inverse, center, middle

### What about rasters?
---
## Rasters
```{r raster-ex, eval=FALSE}

worldclim <- raster::getData("worldclim", var = "bio", res = 10)
r <- worldclim[[2]]

pal <- colorNumeric("RdYlBu", values(r), na.color = "transparent")

leaflet() %>% addTiles() %>%
  addRasterImage(r, colors = pal, opacity = 0.8) %>%
  addLegend(pal = pal, values = values(r), title = "worldclim[[2]]")
```

---

```{r raster-map, out.width='100%', out.height = '100%', echo=FALSE, message=FALSE, warning=FALSE}

worldclim <- raster::getData("worldclim", var = "bio", res = 10)
r <- worldclim[[2]]

pal <- colorNumeric("RdYlBu", values(r), na.color = "transparent")

leaflet() %>% addTiles() %>%
  addRasterImage(r, colors = pal, opacity = 0.8) %>%
  addLegend(pal = pal, values = values(r), title = "worldclim[[2]]")
```
---
class: inverse, center, middle

### OK, so what do I do with this?

---
# Applications of Leaflet

Leaflet maps can be embeded into HTML-based RMarkdown files (including presentations like this one!)
* if you're interested, these slides were made with the [`xaringan` package](https://github.com/yihui/xaringan)

--

You can export Leaflet maps as an image with `mapview` package:
```{r mapview, eval=FALSE}
library(mapview)
mapview::mapshot(fl_map, file = "Florida_Counties.png")
```

--

## SHINY!!

---
class: inverse, center, middle

# Shiny Example

https://github.com/ufrmeetup/gruShiny/tree/master/app3


